### 2022-03-02 index.Rmd
prep for conf slides

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
subtitle: "A Secondary Analysis of SPRINT"
author: "Nicholas Pajewski, PhD"
institute: "Wake Forest School of Medicine"
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

source("packages.R")

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 10, 
                      fig.height = 8)

tar_load(names = c(base_viz_inc,
                   base_viz_eff,
                   long_viz,
                   bp_viz,
                   ndi_baseline))

fig <- left_join(base_viz_inc, base_viz_eff)

```

???

Notes for you go here

---

```{r}

fig$inc_acm[[1]]

# If you wanted to make multiple figs on the same slide
# blank_plot <- ggplot(data.frame()) + aes() + theme_void()
# 
# legend <- get_legend(
#   fig$inc_acm[[1]]$plot + 
#     theme(legend.position = c(.55, .2),
#           legend.direction = 'horizontal')
# )
# 
# plot_grid(
#   # row 1 (legend)
#   legend, 
#   blank_plot,
#   # row 2 (figures)
#   fig$inc_acm[[1]]$plot + theme(legend.position = ''), 
#   fig$eff_acm[[1]],
#   # row 3 (table)
#   fig$inc_acm[[1]]$table, 
#   blank_plot,
#   ncol = 2,
#   align = 'v',
#   axis = 'lr',
#   rel_heights = c(1,10,2),
#   labels = c("","","A","B","",""),
#   label_x = 0.1,
#   label_y = 1.05
# )

```

---