### 2022-03-14 index.Rmd
slides first draft, PK&WC feedback

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
subtitle: "A Secondary Analysis of SPRINT"
author: "Nicholas M. Pajewski, PhD & Byron C. Jaeger, PhD"
institute: "Wake Forest School of Medicine"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

# IN CASE YOU NEED THIS IN PDF: Run this code in console:
# pagedown::chrome_print('index.html')
# But avoid using the pdf version when you present b/c you 
# can access keyboard shortcuts if you present from html.
# e.g., presenter mode by pressing 'p'

options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600)

tar_load(names = c(recoders,
                   base_viz_inc,
                   base_viz_eff,
                   bp_viz,
                   tbl_characteristics,
                   base_ds, 
                   long_ds,
                   long_inf,
                   trt_hetero,
                   base_inf,
                   ndi_baseline))

long_viz <-
  ndi_long_viz(long_inf, trt_hetero, long_ds, recoders, size_text = 4.5)

n_total <- nrow(ndi_baseline)

n_ehr <- sum(ndi_baseline$ehr_ancillary == 'Yes')

fig <- left_join(base_viz_inc, base_viz_eff)

```

## Importance and Background

- SPRINT showed that intensive treatment reduced the risk of cardiovascular and all-cause mortality.

- The trial began in 2010 and the last study closeout visit occurred in 2016.

## Objective and Methods

- We assessed whether the benefit of intensive treatment persisted after the final study closeout visit in 2016.

- We tracked cardiovascular and all-cause mortality events from the US National Death Index from 2016 to 2020.

- In a subset of participants, we examined outpatient systolic blood pressure levels measured in routine clinical practice after the trial.

???

intensive treatment: Systolic blood pressure goal of < 120 mm Hg

legacy effect: the persistence of benefit after the trial

---

```{r fig.width = 10, fig.height = 8}

fig$inc_cvd[[1]] +
  theme(legend.key.size = unit(1.2, 'cm'),    
        legend.key.height = unit(1.2, 'cm'),  
        legend.key.width = unit(1.2, 'cm'),   
        legend.title = element_text(size=16), 
        legend.text = element_text(size=16))

```

---

```{r fig.width = 10, fig.height = 8}

bp_viz$levels + 
  theme(legend.position = 'top',
        legend.key.size = unit(1.2, 'cm'), #change legend key size
        legend.key.height = unit(1.2, 'cm'), #change legend key height
        legend.key.width = unit(1.2, 'cm'), #change legend key width
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=16))

```

---

```{r fig.width = 10, fig.height = 8}

long_viz$cvd$cohort

```

---

## Key findings

- Intensive treatment produced beneficial effects on mortality during the trial.

- The benefits associated with intensive treatment quickly attenuated as systolic blood pressure levels increased among participants who underwent intensive treatment after the trial.

- There was no evidence of sustained benefits after discontinuing the intervention protocol.

## Conclusion

Maintaining intensive BP targets throughout adulthood is likely an essential component of long-term CVD risk management.