### 2024-07-05 index.qmd
updated ndi

---
title: "Longer Term All-Cause and Cardiovascular Mortality With Intensive Blood Pressure Control"
subtitle: "A Secondary Analysis of a Randomized Clinical Trial"
author: "Byron C. Jaeger, PhD"
institute: "Wake Forest University School of Medicine"
format: 
  revealjs:
    slide-number: true
    footer: Slides available at
    width: 1280
    height: 720
    fig-dpi: 300
    fig-width: 8
    fig-height: 6
    fig-align: "center"
    message: false
    bibliography: doc/refs.bib
    reference-location: document
knitr:
  opts_chunk:
    out.width: "100%"
---

```{r}

withr::with_dir(
  here::here(),
  code = {
    source("packages.R")
    targets::tar_load(
      names = c(long_viz,
                bp_viz,
                base_viz_inc,
                base_sub_overall)
    )
  }
)

ltd_2020 <- read_csv("T:/sprint/npajewski/NDI/Data/longterm_death.csv")
ltd_2023 <- read_csv("T:/sprint/npajewski/NDI/Data/longterm_death_2023.csv")

n_cvd_2020 <- sum(ltd_2020$cvd_event_cr == 1, na.rm = TRUE)
n_cvd_2023 <- sum(ltd_2023$cvd_event_cr == 1, na.rm = TRUE)

n_acm_2020 <- sum(ltd_2020$acm_event, na.rm = TRUE)
n_acm_2023 <- sum(ltd_2023$acm_event, na.rm = TRUE)

theme_fig <- function(x, ...){
  theme_bw() +
    theme(text = element_text(size = 18),
          axis.text = element_text(size = 18),
          axis.title = element_text(size = 18),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          legend.text = element_text(size = 18),
          legend.title = element_text(size = 18),
          legend.background = element_blank(),
          ...)
}

theme_tbl <- function(x, ...) {
  theme_void() +
    theme(text = element_text(size = 18),
          plot.caption = element_text(hjust = 0, size = 2),
          plot.title.position = "plot",
          plot.caption.position =  "plot",
          ...)
}

cols_bg <- viridis(3)
cols_tx <- pal_lancet()(2)

```

## Overview

- Background: SPRINT

- Why is longer term follow-up relevant?

- How did SPRINT do longer term follow-up on *all-cause and CVD mortality*?

- How did we analyze the longer term follow-up?

- Results

- Conclusion

## SPRINT

Two treatments with different systolic blood pressure (SBP) targets:

- *Standard*: SBP target of < 140 mm Hg
- *Intensive*: SBP target of < 120 mm Hg

Participants:

- aged 50 years or older 
- with hypertension and increased cardiovascular risk
- without diabetes or history of stroke

## Why do longer term follow-up?

::: {.column width="39%"}

- Investigate the legacy effect of intensive treatment

- Does targeting SBP < 120 mm Hg have long term benefits or harms compared to targeting SBP < 140 mm Hg?

:::

::: {.column width="59%"}

```{r}

ndi_data <- base_sub_overall$overall

fit_cr <- cuminc(ftime   = ndi_data$acm_years,
                 fstatus = factor(ndi_data$cvd_event_cr,
                                  levels = c(0, 1, 2),
                                  labels = c("- Censored",
                                             "- CVD mortality",
                                             "- Non-CVD mortality")),
                 group   = ndi_data$treatment,
                 cencode = '- Censored')

data_cr <-
  map_dfr(.x = fit_cr[1:4],
          .f = as_tibble,
          .id = 'fit') |>
  separate(fit, into = c('treatment','event'), sep = ' - ') |>
  mutate(
    treatment = factor(
      x = treatment,
      levels = c("Standard", "Intensive")
    ),
    lwr = est - 1.96 * sqrt(var),
    upr = est + 1.96 * sqrt(var)
  )

ymax <- 0.10

fig_cr <- ggplot(filter(data_cr, time < 4.48)) +
  aes(x = time,
      y = est,
      col = treatment,
      group = interaction(treatment, event)) +
  geom_line(linewidth = 1.5) +
  geom_ribbon(
    aes(x = time,
        ymin = est - sqrt(var) * 1.96,
        ymax = est + sqrt(var) * 1.96,
        fill = treatment,
        group = interaction(treatment, event)),
    alpha = 0.10,
    inherit.aes = FALSE
  ) +
  theme_fig(legend.position = 'inside',
            legend.position.inside = c(0.15, 0.75)) +
  labs(linetype = 'Event',
       color = 'Treatment',
       fill = 'Treatment') +
  ylab('Cumulative incidence') +
  xlab('Years since randomization') +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0, ymax, by = 0.05),
                     limits = c(0, ymax)) +
  scale_x_continuous(breaks = c(0:12),
                     expand = c(0, 0, 0, 0),
                     limits = c(0, 13)) +
  scale_color_manual(values = cols_tx) +
  scale_fill_manual(values = cols_tx) + 
  annotate(geom = "text", hjust = 0, size = 7,
           x = 5, y = .045, label = "Non-CVD mortality") + 
  annotate(geom = "text", hjust = 0, size = 7,
           x = 5, y = 0.0175, 
           label = "CVD mortality") + 
  annotate(geom = "text", hjust = 0, size = 7,
           x = 11, y = 0.07, 
           label = "What\nhappens\nhere??")

fig_cr

```

:::

## Longer term all-cause/CVD mortality

::: {.column width="49%"}

- We assessed all-cause^[Deaths were treated as confirmed if they were a Class 1 match, or a Class 2, 3, or 4 match with a probabilistic score above cutoffs recommended by the NDI.] and cardiovascular disease^[Cardiovascular mortality for NDI-based follow-up used the NDI Plus System, which automatically identifies underlying causes of death from death certificates, including conversion to International Statistical Classification of Diseases and Related Health Problems, Tenth Revision (ICD-10) codes. We defined CVD mortality as any death containing the ICD-10 codes of I00 to I99] (CVD) mortality post-trial via the US National Death Index (NDI) from 2016-2020 [@jaeger2022longer].

- Among 2944 trial participants,^[We identified 3074 participants with 3 or more electronic health record reports of outpatient blood pressure measurements during the trial. After excluding 130 participants without electronic health record data following July 2016 (ie, conclusion of the trial), a total of 2944 patients were included for the ancillary blood pressure analysis.] post-trial SBP levels were analyzed [@drawz_concordance_2020].

:::

::: {.column width="49%"}
![](img/jama_cardio_cover.png)
:::

## How did we analyze longer term follow-up?

![](img/ph-assumption.gif){.absolute top="120" right="0"}

::: {.column width="50%"}
- assumed treatment group differences would not be constant over time.

- split each participant’s follow-up time into non-overlapping trial and observational phases

- estimated regression coefficients for intensive treatment separately during each phase.

:::

::: {.column width="49%"}
:::

# Results

## What's changed?

::: {.column width="49%"}
**2016-2020**

- `r n_acm_2020` all-cause mortality events
- `r n_cvd_2020` CVD mortality events
:::

::: {.column width="49%"}
**2016-2023**

- `r n_acm_2023` all-cause mortality events
- `r n_cvd_2023` CVD mortality events
:::

**Difference**:

- `r n_acm_2023 - n_acm_2020` additional all-cause mortality events
- `r n_cvd_2023 - n_cvd_2020` additional CVD mortality events
- No additional analyses on BP

## Systolic BP levels

::: {.column width="15%"}

SBP levels among the intensive group increased after the trial.

:::

::: {.column width="4%"}

:::

::: {.column width="79%"}
```{r}
#| fig-width: 10

bp_viz$levels + 
  theme_fig(legend.position = 'top')

```
:::

## Systolic BP difference

::: {.column width="15%"}
By year 10, the estimated difference was about 0
:::

::: {.column width="4%"}

:::

::: {.column width="79%"}
```{r}
#| fig-width: 10

bp_viz$diff + 
  theme_fig(legend.position = 'top') + 
  ylab("Between-group difference in SBP, mm Hg")

```
:::

## All-cause mortality (2016-2020)

::: {.column width="15%"}
Our original analysis showed the benefit for all-cause mortality attenuated quickly
:::

::: {.column width="4%"}
:::

::: {.column width="79%"}

```{r}
#| fig-width: 10

base_viz_inc$inc_acm[[1]]$plot + 
  annotate(geom = 'rect',
           xmin = 9.8,
           ymin = .18,
           xmax = Inf,
           ymax = .35,
           fill = 'black',
           alpha = 1) + 
  theme_fig(legend.position = 'inside',
            legend.position.inside = c(0.75, 0.20))

```
:::

## All-cause mortality (2016-2023)

::: {.column width="15%"}
This update shows an unexpected separation after several years of overlap
:::

::: {.column width="4%"}
:::

::: {.column width="79%"}

```{r}
#| fig-width: 10

base_viz_inc$inc_acm[[1]]$plot + 
  theme_fig(legend.position = 'inside',
            legend.position.inside = c(0.75, 0.20))

```

:::

## CVD mortality (2016-2020)

::: {.column width="15%"}
Our original analysis showed benefit for CVD mortality attenuated.
:::

::: {.column width="4%"}
:::

::: {.column width="79%"}

```{r}
#| fig-width: 10

base_viz_inc$inc_cvd[[1]] + 
  annotate(geom = 'rect',
           xmin = 9.8,
           ymin = .02,
           xmax = 13.2,
           ymax = .25,
           fill = 'black',
           alpha = 1) + 
  theme_fig(legend.position = 'inside',
            legend.position.inside = c(0.5, 0.69))

```

:::

## CVD mortality (2016-2023)

::: {.column width="15%"}
This update is a bit of a head-scratcher.
:::

::: {.column width="4%"}
:::

::: {.column width="79%"}
```{r}
#| fig-width: 10

base_viz_inc$inc_cvd[[1]] +
  theme_fig(legend.position = 'inside',
            legend.position.inside = c(0.5, 0.69))

```
:::

## All-cause mortalty (time-split analysis)

```{r}
#| fig-width: 14

library(cowplot)

plot_grid(
  long_viz$acm$trial,
  long_viz$acm$cohort,
  nrow = 1
)

```

## CVD mortalty (time-split analysis)

```{r}
#| fig-width: 14

plot_grid(
  long_viz$cvd$trial,
  long_viz$cvd$cohort,
  nrow = 1
)

```

## References

::: {#refs}
:::